---
title: "R Notebook"
output: html_notebook
---

we want ot develop a model to predict whether a given car gets high or low gas mileage based on Auto data set.

```{r}
summary(Auto)
```
```{r}
pairs(Auto)
```
almost every predictor seem to have correlation with mpg but they also have correlation with each other.
we format the mpg into a binary nominal variable by giving it a value of 1 whenever it has a value above median and 0 whenever it has a value below median
```{r}
Auto$mpg01 <- ifelse(Auto$mpg>median(Auto$mpg),1,0)
Auto$mpg
Auto$mpg01
median(Auto$mpg)
```
now we can fit a classification model on our data.
```{r}
boxplot(Auto,horizontal = TRUE)
```

```{r}
for (i in 1:ncol(Auto)){
    boxplot(Auto[,i],xlab = names(Auto)[i],ylab="value")
}
```
we split data into training and test sets.
```{r}
set.seed(1)
s <- sample.int(397,317,replace = FALSE)
Auto.train <- Auto[s,]
Auto.test <-Auto[-s,]
dim(Auto.test)
```

```{r}
library(MASS)
alda <- lda(mpg01~ acceleration + weight + horsepower + displacement, data= Auto.train)
alda
```
```{r}
alda.pred <- predict(alda,newdata = Auto.test)
alda.pred
```

```{r}
mean(alda.pred$class == Auto.test$mpg01)
```
```{r}
table(alda.pred$class,Auto.test$mpg01)
```
seems like only 9 out of 80 test data were predicted uncorrectly and we got a pretty low test error using lda model.

```{r}
aqda <- qda(mpg01 ~ acceleration + weight + horsepower + displacement, data= Auto.train)
aqda
```
```{r}
aqda.pred <- predict(aqda,newdata = Auto.test)
aqda.pred
```

```{r}
mean(aqda.pred$class == Auto.test$mpg01)
```
```{r}
table(predicted= aqda.pred$class,actual = Auto.test$mpg01)
```
qda test results seem similar to lda model except that error model a little toward 0 predicted values.

```{r}
autolr <- glm(mpg01 ~  I(weight^2) + I(horsepower^3) +displacement, data= Auto.train,family=binomial)
summary(autolr)
```
```{r}
autolr.prob <- predict(autolr,newdata = Auto.test,type = "response")
autolr.prob
```

```{r}
autolr.pred <- ifelse(autolr.prob>0.5,1,0)
mean(autolr.pred == Auto.test$mpg01)
```
```{r}
table(predicted= autolr.pred,actual = Auto.test$mpg01)
```
performed a little worse than lda or qda.

```{r}
library(class)
attach(Auto)
set.seed(1)
```
```{r}
Auto.clean <- na.omit(Auto)
dim(Auto.clean)
dim(Auto)
```
```{r}
392*(8/10)
```
```{r}
s<- sample(392,313)
```
```{r}
train.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[s,]
test.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[-s,]
train.y <- Auto.clean$mpg01[s]
autoknn <- knn(train.knn,test.knn,train.y,k=1,)
table(predicted = autoknn,actual = Auto.clean$mpg01[-s])
```
even with k=1 results seem ok
let's try bigger k values

```{r}
train.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[s,]
test.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[-s,]
train.y <- Auto.clean$mpg01[s]
autoknn <- knn(train.knn,test.knn,train.y,k=2)
table(predicted = autoknn,actual = Auto.clean$mpg01[-s])
```

```{r}
train.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[s,]
test.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[-s,]
train.y <- Auto.clean$mpg01[s]
autoknn <- knn(train.knn,test.knn,train.y,k=3)
table(predicted = autoknn,actual = Auto.clean$mpg01[-s])
```
```{r}
train.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[s,]
test.knn <- cbind(Auto.clean$weight,Auto.clean$horsepower)[-s,]
train.y <- Auto.clean$mpg01[s]
autoknn <- knn(train.knn,test.knn,train.y,k=5,)
table(predicted = autoknn,actual = Auto.clean$mpg01[-s])
```
k=3 and k=5 resulted in same cm, but both are better than smaller k values.
