---
title: "R Notebook"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
```{r}
library(ISLR2)
names(Smarket)
```
```{r}
dim(Smarket)
```
```{r}
summary(Smarket)
```
```{r}
pairs(Smarket)
```
```{r}
cor(Smarket[, -9])
```

```{r}
attach(Smarket)
```
```{r}
plot(Volume)
```

```{r}
glm.fits <- glm(
  Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Volume,
  data=Smarket, family = binomial
)
summary(glm.fits)
```
```{r}
coef(glm.fits)
```
```{r}
summary(glm.fits)$coef
```

```{r}
summary(glm.fits)$coef[, 4]
```
```{r}
glm.probs <- predict(glm.fits, type = "response")
```
```{r}
glm.probs[1:10]
```
```{r}
contrasts(Direction)
```

```{r}
glm.pred <- rep("Down", 1250)
```
```{r}
glm.pred[glm.probs > .5] = "Up"
```

```{r}
table(glm.pred, Direction)
```
```{r}
(507 + 145) / 1250
```

```{r}
mean(glm.pred == Direction)
```

```{r}
train <- (Year < 2005)
```

```{r}
Smarket.2005 <- Smarket[!train, ]
```

```{r}
dim(Smarket.2005)
```
```{r}
Direction.2005 <- Direction[!train]
```
```{r}
glm.fits <- glm(
  Direction ~ Lag1 + Lag2 +Lag3 +Lag4 +Lag5 + + Volume, 
  data = Smarket , family = binomial , subset = train
)
```
```{r}
glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
```
```{r}
glm.pred<- rep("Down", 252)
glm.pred[glm.probs > .5] <- "up"
table(glm.pred, Direction.2005)
```
```{r}
mean(glm.pred == Direction.2005)
```
```{r}
mean(glm.pred != Direction.2005)
```
```{r}
glm.fits <- glm(Direction ~ Lag1+ Lag2, data= Smarket,
                family = binomial , subset = train)
```

```{r}
glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
```
```{r}
glm.pred <-rep("Down", 252)
glm.pred[glm.probs > .5] <- "Up"
table(glm.pred, Direction.2005)
```
```{r}
mean(glm.pred == Direction.2005)
```
```{r}
106 / (106 + 76)
```
```{r}
predict(glm.fits, 
        newdata = data.frame(Lag1 = c(1.2,1.5), lag2 = c(1.1, -0.8)),
        type = "response")
```

```{r}
library(MASS)
lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket,
               subset = train)
lda.fit
```
```{r}
lda.pred <- predict(lda.fit,Smarket.2005)
names(lda.pred)
```
```{r}
lda.class <- lda.pred$class
table(lda.class, Direction.2005)
```

```{r}
mean(lda.class == Direction.2005)
```
```{r}
sum(lda.pred$posterior[,1]>=.5)
```
```{r}
sum(lda.pred$posterior[,1] < .5)
```

```{r}
lda.pred$posterior[1:20,1]
lda.class[1:20]
```

Quadratic Discriminant Analysis

```{r}
qda.fit <- qda(Direction ~ Lag1 + Lag2, data = Smarket, subset = train)
qda.fit
```

```{r}
qda.class <- predict(qda.fit, Smarket.2005)$class
table(qda.class, Direction.2005)
```
```{r}
mean(qda.class == Direction.2005)
```

Naive bayes this time

```{r}
library(e1071)
nb.fit <- naiveBayes(Direction ~ Lag1 + Lag2, data = Smarket, subset = train)
nb.fit
```

```{r}
library(class)
train.X <- cbind(Lag1, Lag2)[train,]
test.X <- cbind(Lag1, Lag2)[!train,]
train.Direction<- Direction[train] 
```

```{r}
set.seed(1)
knn.pred <- knn(train.X,test.X, train.Direction, k = 1)
table(knn.pred, Direction.2005)
```

```{r}
set.seed(1)
knn.pred <- knn(train.X,test.X, train.Direction, k = 3)
table(knn.pred, Direction.2005)
mean(knn.pred == Direction.2005)
```

```{r}
dim(Caravan)
attach(Caravan)
summary(Purchase)
```
```{r}
standardized.X <- scale(Caravan[, -86])
var(Caravan[, 1])
var(Caravan[, 2])
var(standardized.X[, 2])
var(standardized.X[, 2])
```

```{r}
test <- 1:1000
train.X <- standardized.X[-test, ]
test.X <- standardized.X[test, ]
train.Y <- Purchase[-test]
```
```{r}
test.Y <- Purchase[test]
set.seed(1)
knn.pred <- knn(train.X, test.X, train.Y, k = 1)
mean(test.Y != knn.pred)
mean(test.Y != "No")
```

```{r}
table(knn.pred, test.Y)
9/(68+9)
```
though knn with k = 1 is not a very good model most of the time, it turns out it is far better than random guessing in this case(9/77 times it predicter yes the people actually purchased it.)

```{r}
knn.pred <- knn(train.X, test.X, train.Y, k = 3)
table(knn.pred, test.Y)
```
and 19% success rate with k = 3 
```{r}
5/26
knn.pred <-  knn(train.X, test.X, train.Y, k = 5)
table(knn.pred, test.Y)
```
```{r}
4/15
```

```{r}
glm.fits <- glm(Purchase ~ ., data = Caravan, family = binomial, subset = -test)
glm.probs <- predict(glm.fits, Caravan[test,], type="response")
glm.pred <- rep("No", 1000)
glm.pred[glm.probs> 0.5] <- "Yes"
table(glm.pred, test.Y)
```

```{r}
 glm.pred <- rep("No", 1000)
glm.pred[glm.probs > .25] <- "Yes"
table(glm.pred, test.Y)
```
```{r}
11/(22+11)
```
it's right about them when it says yes 33% of the times.
a pretty good model.




#-poisson regression for bikeshare data
```{r}
attach(Bikeshare)
dim(Bikeshare)
names(Bikeshare)
```
```{r}
mod.lm <- lm(
  bikers ~ mnth + hr + workingday + temp + weathersit,
  data = Bikeshare
)
summary(mod.lm)
```

```{r}

```







