this is answers to s1 of ISLR book exercises


```{r}
head(Auto)
```
```{r}
Alm <- lm(mpg ~ horsepower, data = Auto)
summary(Alm)
```
shows beta0 of 40 and beta 1 of .16 which shows moderate relationship between two variables but a very small p-value so we are sure that the relationship do exist and we did not get this estimates on random chance.
the relationship is negative so when horsepower increases we expect the mpg to decrease.

```{r}
predict(Alm, newdata = data.frame(horsepower = 98))
```
using this estimates of beta0 and beta1 the mpg predicted for horsepower of 98 is 24.46708.

```{r}
predict(Alm, newdata = data.frame(horsepower = 98), interval = "prediction", level= 0.95)
```
so the prediction interval is (14.8094, 34.12476)
```{r}
predict(Alm, newdata = data.frame(horsepower = 98), interval = "confidence", level= 0.95)
```
as expected the CI is much smaller than PI because it is talking about interval of mean of predictions.
----------------------------------------------------------------------------------------------
----------------------------------------------------------------------------------------------

```{r}
pairs(Auto)
```
```{r}
cormatrix = cor(Auto[,-which(names(Auto)=="name")])
cormatrix
```

seems like there is correlations between most of predictors and mpg except name.
also there are big correlations between the predictors too which does not matter if we are using multiple linear regression.
```{r}
alm = lm(mpg ~ .-name, data= Auto)
summary(alm)
```
as the correlations suggested there is relationship between mpg and most of the predictors.
but acceleration,cylinders and horsepower have a pretty big p-values so we cannot be sure about a meaningful relationship between them and mpg.
```{r}
plot(alm)
```
in the residual vs fitted plot we can see that residuals aren't constant and the residuals have a funnel shape.
we also see an very high leverage point and some high leverage points in residuals vs leverage plot
```{r}
imlm = lm(mpg ~ . -name + year:origin + origin:weight + origin:displacement + year:weight + year:displacement + weight:displacement,data=Auto )
summary(imlm)
```
```{r}
plot(imlm)
```
seems like there is relationship between mpg and displacement:weight interaction.
but not other significant p-value in interactoin terms.
```{r}
tlm <- lm(mpg ~ . - name + poly(year,3) + poly(weight,3) ,data=Auto)
summary(tlm)
```
```{r}
tlm <- lm(mpg ~ . - name + log(year) + log(weight) + log(horsepower) ,data=Auto)
summary(tlm)
```
found lots of relationships between transformations of predictors too.
we can combine the all important predictors and transformations to build a more accurate model in the end.

```{r}
Carseats
```
```{r}
pairs(Carseats )
```

```{r}
Carseatsmlm <- lm(Sales ~ Price + Urban + US, data=Carseats)
summary(Carseatsmlm)
```
okay seems like we can be sure about relationship between Price and USYes predictors but UrbanYes has a very high p-value we can almost be sure there's no relationship.
relationship with price is negative so we can conclude that if we increase carseat prices the sales would decrease.
coefficient for UrbanYes was estimated slightly negative but since the p-value is very high we cannot inference anything from this coefficient.
and USYes also has a high positive coefficient which indicates that if the store is in the US the sales would increase.

the equation form is like
Sales = (-0.054459)xPrice + (-0.021916) + 1.200573
if the Urban and US are yes
and like
Sales = (-0.054459)xPrice
if both are no.

```{r}
Carseatsmlm2 <- lm(Sales ~ Price + US, data=Carseats)
summary(Carseatsmlm2)
```
```{r}
confint(Carseatsmlm2,level=0.95)
predicint(Carseatsmlm2,)
```
```{r}
plot(Carseatsmlm2)
```
there is some high leverage points but I don't see any outliers.
--------------------------------------------------------------------
--------------------------------------------------------------------
11.
we generate a predictor x and a response y as follows
```{r}
set.seed(11)
x <- rnorm(100)
y <- 2 * x +rnorm(100)
y
```
we perform a simple linear regression without an intercept with
```{r}
ylm <- lm(y~x+0)
summary(ylm)
```
the coefficient was estimated 2 with a very small p-value that shows that x and y have a positive relationship.
t-value of 18.63 is very high and easily rejects null hypothesis of betha0 = 0
```{r}
confint(ylm)
```
the 95% confident interval of intercept is also arount 1.8 and 2.2 

now we perform a simple linear regression
```{r}
yslm <- lm(y ~ x)
summary(yslm)
```
the p-value of intercept is high so we cannot reject the null-hypothesis of beta0 = 0
the standard error also didn't change between simple linear regression and linear regression without intercept so we can say that intercept didn't have any impact on improving our model.

q11)d)
```{r}
summary(ylm)
```
we compute the t-statistics with formula given for the regression without intercept
```{r}
n <- length(x)
sum_xy <- sum(x*y)
sumx2 <- sum(x^2)
sumy2 <- sum(y^2)

t_formula <- (sqrt(n-1)*sum_xy) / sqrt(sumx2 * sumy2 - sum_xy^2)
```
now we compute t-statistic using regression without intercept
```{r}
t_regresstion <- summary(ylm)$coefficients[1,"t value"]
```
and compare them:
```{r}
t_formula
t_regresstion
identical(round(t_formula, 10), round(t_regresstion, 10))
```
we see that the both numbers are the same so the t-statistic implemented in R and the given formula are the same.

q11)e)

by changing the places of y and x in the formula given, the result would not change so t-statistic for regression without intercept of y into x or x into y would be the same.

q11)f)
we perform the regression for y into x
```{r}
xlm <- lm(x~y+0)
```
```{r}
summary(xlm)$coefficients[1,"t value"]
```
the t-value is still the same.
---------------------------------------------------------------------------------
---------------------------------------------------------------------------------
q12)a)
in regression without intercept, if sum of all x^2 is equal to sum of all y^2 the coefficient of both regression from y into x and x into y will be the same.

b)
```{r}
x1 <- rnorm(100)
y1 <- x1^2 - 20 +rnorm(100)
yinx1 <- lm(y1 ~ x1+0)
summary(yinx1)$coefficients[1]
xiny1 <- lm(x1 ~ y1+0)
summary(xiny1)$coefficients[1]
```
```{r}
x2 <- rnorm(100)
y2 <- x2 * rnorm(100)
yinx2 <- lm(y2 ~ x2+0)
summary(yinx2)$coefficients[1]
xiny2 <- lm(x2 ~ y2+0)
summary(xiny2)$coefficients[1]
```
they're haven't become exactly the same but they are very close because they have close distribution and std.

q13)a)
```{r}
set.seed(1)
x <- rnorm(100)
eps <- rnorm(100,0,0.25)
y <- -1+ 0.5*x +eps
length(y)
yixlm <- lm(y ~x)
summary(yixlm)
```
the estimated betas are very close to the given intercept and coef while creating the y.

scatter plot displaying relationship of y and x:
```{r}
plot(x,y)
abline(yixlm,col="blue")
legend("bottomright",legend = c("actual","estimate"),fill = c("gray","blue"))
```
```{r}
yixplm <- lm(y ~ poly(x,degree=2))
summary(yixplm)
```
well the RSE and R-squared slightly improved but p_value for third coefficient is very high and indicate no relationship with y. and perhabs if we test on test data it would even give worse resaults.


now we decrease the error
```{r}
x <- rnorm(100)
eps <- rnorm(100,0,0.1)
y <- -1+ 0.5*x +eps
length(y)
yixlm2 <- lm(y ~x)
summary(yixlm2)
```

```{r}
plot(x,y)
abline(yixlm2,col="blue")
legend("bottomright",legend = c("actual","estimate"),fill = c("gray","blue"))
```
the regression line fits the data much better than the original data and thus we see less error.


we generate more noisier model:

q13)a)
```{r}
x <- rnorm(100)
eps <- rnorm(100,0,0.35)
y <- -1+ 0.5*x +eps
length(y)
yixlm3 <- lm(y ~x)
summary(yixlm3)
```
the estimated betas are very close to the given intercept and coef while creating the y.

scatter plot displaying relationship of y and x:
```{r}
plot(x,y)
abline(yixlm3,col="blue")
legend("bottomright",legend = c("actual","estimate"),fill = c("gray","blue"))
```
```{r}
predict(yixlm,newdata = data.frame(x = 5),interval = "prediction")
predict(yixlm2,newdata = data.frame(x = 5),interval = "prediction")
predict(yixlm3,newdata = data.frame(x = 5),interval = "prediction")
predict(yixlm,newdata = data.frame(x = 5),interval = "confidence")
predict(yixlm2,newdata = data.frame(x = 5),interval = "confidence")
predict(yixlm3,newdata = data.frame(x = 5),interval = "confidence")


```
the interval moves for less and more variable data but the interval is about the same size.

q14)
```{r}
set.seed(1)
x1 <- runif(100)
x2 <- 0.5 * x1 + rnorm(100)/10
y <- 2 + 2 * x1 + 0.3 * x2 + rnorm(100)
```
we made data with beta0 equal to 2, beta1 equal to 2,beta2 equal to 0.3 and a standard normal distributed error.
there is close to one correlation between x1 and x2.
```{r}
plot(x1,x2)
cor(x1,x2)
```
```{r}
pred <- lm(y ~ x1 + x2)
summary(pred)
```
model is not sure about x2 having a relationship with y
and the estimates for beta1 and beta2 are very far from reallity.
seems like although we chose the right model to describe the linear relationship we have, but having correlated predictors ruined the model.
```{r}
lmx1 <- lm(y~x1)
summary(lmx1)
```
the coefficient for this one seems right.

```{r}
lmx2 <-lm(y ~ x2)
summary(lmx2)
```
Now suppose we obtain one additional observation, which was
unfortunately mismeasured.
```{r}
x1 <- c(x1, 0.1)
x2 <- c(x2, 0.8)
y <- c(y, 6)
```
```{r}
plot(x1,x2)
cor(x1,x2)
plot(y)
```
```{r}
pred <- lm(y ~ x1 + x2)
summary(pred)
```
```{r}
lmx1 <- lm(y~x1)
summary(lmx1)
```


```{r}
lmx2 <-lm(y ~ x2)
summary(lmx2)
```
the added point is a high leverage point in x2
also have a higher value than most of the point in y so can be considered on of outliers
though it is an outlier in x2 but it has affected the lmx1 model more, because x1 and x2 are correlated.

q15)a)
```{r}
library(ISLR2)
attach(Boston)
crim.zn <- lm(crim ~ zn)
crim.indus <- lm(crim ~ indus)
crim.chas <- lm(crim ~ chas)
crim.nox <- lm(crim ~ nox)
crim.rm <- lm(crim ~ rm)
crim.age <- lm(crim ~ age)
crim.dis <- lm(crim ~ dis)
crim.rad <- lm(crim ~ rad)
crim.tax <- lm(crim ~ tax)
crim.ptratio <- lm(crim ~ ptratio)
crim.lstat <- lm(crim ~ lstat)
crim.medv <- lm(crim ~ medv)
summary(crim.zn)
```
```{r}
summary(crim.indus)
```
```{r}
summary(crim.chas)
```
```{r}
summary(crim.zn)
```
```{r}
summary(crim.nox)
```
```{r}
summary(crim.nox)
```
```{r}
summary(crim.rm)
```
```{r}
summary(crim.age)
```
```{r}
summary(crim.dis)
```
```{r}
summary(crim.rad)
```
```{r}
summary(crim.tax)
```
```{r}
summary(crim.ptratio)
```
```{r}
summary(crim.lstat)
```
```{r}
summary(crim.medv)
```
seems like simple linear regression found all of predictors having statistically significant correlation with per capita crime rate of town.
```{r}
crim.mlm <- lm(crim ~ .,data = Boston)
summary(crim.mlm)
```
in multiple linear regression setting, only dis,rad and medv have a very low p-value but and we can definitely regect null hypothesis 
there are other love p-values too and assuming under 0.05 rejection threshold we can reject null hypothesis for zn too.

```{r}
smlcfs <- c(crim.zn$coefficients[2], crim.indus$coefficients[2], crim.chas$coefficients[2], crim.nox$coefficients[2], crim.rm$coefficients[2], crim.age$coefficients[2], crim.dis$coefficients[2], crim.rad$coefficients[2], crim.tax$coefficients[2], crim.ptratio$coefficients[2], crim.lstat$coefficients[2], crim.medv$coefficients[2])
plot(crim.mlm$coefficients[-1],smlcfs)
```
two model coefficients seem very different from each other.
probably because multiple linear regression cancels out correlations between predictors.
now we fit linear model with non-linear predictors
```{r}
crim.nldis <- lm(crim ~ poly(dis,3))
crim.nlrad <- lm(crim ~ poly(dis,3))
crim.nlzn <- lm(crim ~ poly(zn,3))
crim.nlmedv <- lm(crim ~ poly(medv,3))
summary(crim.nldis)
```
```{r}
summary(crim.nlrad)
```
```{r}
summary(crim.nlzn)
```
```{r}
summary(crim.nlmedv)
```
all of the predictors we used seem to have non-linear relations too.
```{r}

```


